---
title: "Ch2_results"
format: html
editor: visual
---

```{r setup chunk}
#| message: false
#| echo: false
#| warning: false
library(tidyverse)
library(tidylog)
library(nlme)
library(MuMIn)
library(DHARMa)
library(tictoc)
library(ggridges)
library(emmeans)
library(multcomp)
library(conflicted)
library(performance)
library(see)
conflict_prefer_all("tidylog", quiet = TRUE)

###=========================================###
###Loop to run SES~elevation for all traits####
###=====C5 NULL MODEL, pool = entire========####
###=========================================###
#import SES data
cell_ses <- read.csv("All_data/comm_assembly_results/SES_RQ_weighted_cells_C5_entire.csv", row.names = 1) |> 
  rename(Cell_ID = cellref) 

#import microenvironmental data containing x and y coordinates
env <- read.csv("All_data/clean_data/Environmental data/All_Sites_Environmental_Data.csv") |> 
  #variables we are interested in
  select(Cell_ID:row) |> 
  #add elevation variables
  mutate(elevation = case_when(site == "GG" ~ 2000, 
                               site == "WH" ~ 2500, 
                               site == "BK" ~ 3000,
                               .default = NA))

##Combine SES and environmental data
comb <- env |> 
  #join, one row in env matches many rows in cell_ses due to it containing ses of different traits
  inner_join(cell_ses, by = "Cell_ID", relationship = "one-to-many") |>
  mutate(ncolumn = match(column, LETTERS[1:8]), 
         grid = paste0(site, grid)) |> #each grid must have a unique id 
  rename(x_coord = ncolumn, 
         y_coord = row)
  
#These variables have the best diagnostics
#LA and SLA still have TERRIBLE DIAGNOSTICS
traitlist <- c("log_Height", "log_LDMC", "log_LA", "SLA")
#lists to store results in
SES_ele_summary <- vector(mode= "list", length = length(traitlist))
names(SES_ele_summary) = traitlist

SES_ele_Rsq<- vector(mode= "list", length = length(traitlist))
names(SES_ele_Rsq) = traitlist

SES_ele_anova<- vector(mode= "list", length = length(traitlist))
names(SES_ele_anova) = traitlist

SES_ele_cld<- vector(mode= "list", length = length(traitlist))
names(SES_ele_cld) = traitlist

for (t in 1:length(traitlist)) {
  modeldat <-  comb |> 
    filter(trait == traitlist[t]) |> 
    mutate(elevation = as.factor(elevation), 
           grid = as.factor(grid)) |> 
    drop_na()
  
  
  model<- lme(SES ~ elevation ,
                  random = ~1|grid, 
                  correlation = corSpher(form = ~ x_coord + y_coord|grid, nugget = TRUE), #spherical structure
                  data = modeldat) #only gaussian family possible
  
  em_model <- emmeans(model, specs = "elevation", type = "response")
  comp_letters <-cld(em_model, Letters = letters, adjust = "Tukey", sort = FALSE)

  #Also save results in a list object so we can call it with quarto
  SES_ele_summary[[t]] <- summary(model)
  SES_ele_Rsq[[t]] <- r.squaredGLMM(model)
  SES_ele_anova[[t]] <- anova(model)
  SES_ele_cld[[t]] <- comp_letters
  
}

###Using the logged traits are generally better for model diagnostics than the raw traits
##BUT the leaf area diagnostics are still very bad




###===========================================================###
###Loop to run SES~microenvironmental variables for all traits####
###=============C5 NULL MODEL, pool = site====================####
###===========================================================###
#import ses
cell_ses_poolsite <- read.csv("All_data/comm_assembly_results/SES_RQ_weighted_cells_C5_poolsite.csv", row.names = 1) |> 
  rename(Cell_ID = cellref) 

#import microenvironmental data
env2 <- read.csv("All_data/clean_data/Environmental data/All_Sites_Environmental_Data.csv") |> 
  #variables we are interested in
  select(Cell_ID:row, rock_cover, northness, soil_moisture_adj_campaign2, 
         soil_depth_CV, mean_soil_depth, slope_height) |> 
  #add elevation variables
  mutate(elevation = case_when(site == "GG" ~ 2000, 
                               site == "WH" ~ 2500, 
                               site == "BK" ~ 3000,
                               .default = NA))

#import remote sensing derived variables
rms <- read.csv("All_data/clean_data/Environmental data/Zonal_stats_all.csv") |> 
  select(CELL_ID, STD) |> 
  rename(Cell_ID = CELL_ID)

#import interpolated microclimate indices
micro_idw <- read.csv("All_data/clean_data/Environmental data/Imke_microclimate_indices_idw_interpolated.csv", row.names = 1)


##Combine SES and environmental data
comb2 <- env2 |> 
  #join to microclimate indices |> 
  full_join(micro_idw, by = "Cell_ID") |> 
  #join to remote sensing data |> 
  full_join(rms, by = "Cell_ID") |> 
  #join, one row in env matches many rows in cell_ses due to it containing ses of different traits
  full_join(cell_ses_poolsite, by = "Cell_ID", relationship = "one-to-many") |>
  mutate(ncolumn = match(column, LETTERS[1:8]), 
         grid = paste0(site, grid)) |> #each grid must have a unique id 
  rename(x_coord = ncolumn, 
         y_coord = row)


###Loop starts here
#These variables have the best diagnostics after comparing diagnostoc plots of log and raw variables
traitlist <- c("log_LDMC", "Height_cm", "SLA", "Leaf_area_mm2")
#lists to store results in
SES_microenv_summary <- vector(mode= "list", length = length(traitlist))
names(SES_microenv_summary) = traitlist

SES_microenv_Rsq<- vector(mode= "list", length = length(traitlist))
names(SES_microenv_Rsq) = traitlist

SES_microenv_anova<- vector(mode= "list", length = length(traitlist))
names(SES_microenv_anova) = traitlist

for (t in 1:length(traitlist)) {
  modeldat <-  comb2 |> 
    filter(trait == traitlist[t]) |> 
    mutate(elevation = as.factor(elevation), 
           grid = as.factor(grid)) |> 
    drop_na()
  
  
  model<- lme(SES ~ mean_T1_growing_season + mean_moist_growing_season + rock_cover+ mean_soil_depth + STD,
              random = ~1|grid, 
              correlation = corSpher(form = ~ x_coord + y_coord|grid, nugget = TRUE), #spherical structure
              data = modeldat) #only gaussian family possible
  

  #Also save results in a list object so we can call it with quarto
  SES_microenv_summary[[t]] <- summary(model)
  SES_microenv_Rsq[[t]] <- r.squaredGLMM(model)
  SES_microenv_anova[[t]] <- anova(model)

}
#plot(checkmodel()) returns the following error: Check it out when you have internet
#Converting missing values (`NA`) into regular values currently not possible for variables of class `NULL`.

```

# Results

`r comb |>  filter(trait == "log_Height") |> drop_na() |> summarise(n = n())` Cells were included in the analysis, with a total species richness of xx. Elevation did not have a consistent effect on the community assembly of different plant traits (Figure 1). Elevation shifted the assembly of plant height and LDMC from environmental filtering at the low elevation, to limiting similarity at the high elevation. Environmental filtering was the dominant assembly mechanism at all elevations for leaf area and SLA, although the filtering effect of SLA was weakest at the mid elevation.

```{r}
#| message: false
#| echo: false
#| label: fig-SES-elevation-ridges
#| fig-cap: "The deviation of trait diversity calculated with Height, SLA, LDMC and leaf area from random assembly (SES)"
#| warning: false


l1 <- c("log_Height" = "SES~of~plant~height", "SLA" = "SES~of~SLA", "log_LDMC" = "SES~of~LDMC", "log_LA" = "SES~of~LA")
ridges_letters <- data.frame(trait = c(rep("log_Height", 3), rep("SLA", 3), rep("log_LDMC", 3),rep("log_LA", 3)), 
                             elevation = as.factor(c(rep(c("2000", "2500", "3000"), 4))),
                             letters = c(SES_ele_cld$log_Height$.group, 
                                         SES_ele_cld$SLA$.group, 
                                         SES_ele_cld$log_LDMC$.group, 
                                         SES_ele_cld$log_LA$.group), 
                             x_pos = c(4.5, 4.5, 4.5,
                                       6,6,6, 
                                       9,9,9,
                                       8.1,8.1,8.1), 
                             emmean = c(SES_ele_cld$log_Height$emmean,
                                        SES_ele_cld$SLA$emmean,
                                        SES_ele_cld$log_LDMC$emmean,
                                        SES_ele_cld$log_LA$emmean), 
                             emmean_SE = c(SES_ele_cld$log_Height$SE,
                                           SES_ele_cld$SLA$SE,
                                           SES_ele_cld$log_LDMC$SE,
                                           SES_ele_cld$log_LA$SE))

ses_ridges <- comb |>
  filter(trait %in% c("log_Height", "SLA", "log_LDMC", "log_LA")) |> 
  mutate(elevation = as.factor(elevation)) |> 
  ggplot(aes(x = SES, y = elevation)) +
  geom_density_ridges(alpha = 0.5) +
  facet_wrap(~trait, labeller = as_labeller(l1, default = label_parsed), scale = "free_x", 
             strip.position = "bottom")+
  labs(x = " ", y = "Elevation (m a.s.l.)") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_text(data = ridges_letters, aes(x = x_pos, y = elevation, label = letters, size = 16))+
  geom_point(
    data = ridges_letters,
    aes(x = emmean, y = as.numeric(elevation)),
    size = 2, inherit.aes = FALSE) +
  geom_errorbarh(
    data = ridges_letters,
    aes(xmin = emmean - emmean_SE,
        xmax = emmean + emmean_SE,
        y = as.numeric(elevation)),
    height = 0.08,
    linewidth = 1, inherit.aes = FALSE)+
  theme_bw() +
  theme(legend.position = "none", axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), strip.text = element_text(size = 18), 
        strip.background = element_blank(),
        strip.placement = "outside", panel.grid = element_blank()) 

ses_ridges
```
Microenvironmental variables....

```{r}
#| label: tbl-gt
#| tbl-cap: Effects of microenvironmental variables on assembly of traits
#| echo: false

library(gt)

microenv_results <- tibble(variable = c("mean soil temperature", "mean soil moisture", "rock cover", "mean soil     depth", "standard deviation of elevation"),
                height_coef = SES_microenv_summary$Height_cm$coefficients$fixed[2:6], 
                height_pval = SES_microenv_anova$Height_cm$`p-value`[2:6], 
                la_coef = SES_microenv_summary$Leaf_area_mm2$coefficients$fixed[2:6], 
                la_pval = SES_microenv_anova$Leaf_area_mm2$`p-value`[2:6],
                ldmc_coef = SES_microenv_summary$log_LDMC$coefficients$fixed[2:6], 
                ldmc_pval = SES_microenv_anova$log_LDMC$`p-value`[2:6],
                sla_coef = SES_microenv_summary$SLA$coefficients$fixed[2:6], 
                sla_pval = SES_microenv_anova$SLA$`p-value`[2:6],)

microenv_results |> 
  gt() |> 
  cols_label(height_coef = "coefficient", 
             height_pval = "p-value", 
             la_coef = "coefficient", 
             la_pval = "p-value", 
             ldmc_coef = "coefficient", 
             ldmc_pval = "p-value", 
             sla_coef = "coefficient", 
             sla_pval = "p-value") |> 
  fmt_number(columns = everything(), decimals = 3) |> 
  tab_spanner(label = "Height", columns = c(height_coef, height_pval)) |>  
  tab_spanner(label = "Leaf area", columns = c(la_coef, la_pval)) |>
  tab_spanner(label = "LDMC", columns = c(ldmc_coef, ldmc_pval)) |>
  tab_spanner(label = "SLA", columns = c(sla_coef, sla_pval))

```

