---
title: "Ch2_results"
format: docx
editor: visual
fig-format: png
fig-dpi: 300
fig-width: 6
---

```{r setup chunk}
#| message: false
#| echo: false
#| warning: false
library(tidyverse)
library(tidylog)
library(nlme)
library(MuMIn)
library(DHARMa)
library(tictoc)
library(ggridges)
library(ggpubr)
library(emmeans)
library(multcomp)
library(conflicted)
library(performance)
library(see)
conflict_prefer_all("tidylog", quiet = TRUE)

###=========================================###
###Loop to run SES~elevation for all traits####
###=====C5 NULL MODEL, pool = entire========####
###=========================================###
#import SES data
cell_ses <- read.csv("All_data/comm_assembly_results/SES_RQ_weighted_cells_C5_entire.csv", row.names = 1) |> 
  rename(Cell_ID = cellref) 

#import microenvironmental data containing x and y coordinates
env <- read.csv("All_data/clean_data/Environmental data/All_Sites_Environmental_Data.csv") |> 
  #variables we are interested in
  select(Cell_ID:row) |> 
  #add elevation variables
  mutate(elevation = case_when(site == "GG" ~ 2000, 
                               site == "WH" ~ 2500, 
                               site == "BK" ~ 3000,
                               .default = NA))

##Combine SES and environmental data
comb <- env |> 
  #join, one row in env matches many rows in cell_ses due to it containing ses of different traits
  inner_join(cell_ses, by = "Cell_ID", relationship = "one-to-many") |>
  mutate(ncolumn = match(column, LETTERS[1:8]), 
         grid = paste0(site, grid)) |> #each grid must have a unique id 
  rename(x_coord = ncolumn, 
         y_coord = row)
  
#These variables have the best diagnostics
#LA and SLA still have TERRIBLE DIAGNOSTICS
traitlist <- c("log_Height", "log_LDMC", "log_LA", "SLA")
#lists to store results in
SES_ele_summary <- vector(mode= "list", length = length(traitlist))
names(SES_ele_summary) = traitlist

SES_ele_Rsq<- vector(mode= "list", length = length(traitlist))
names(SES_ele_Rsq) = traitlist

SES_ele_anova<- vector(mode= "list", length = length(traitlist))
names(SES_ele_anova) = traitlist

SES_ele_cld<- vector(mode= "list", length = length(traitlist))
names(SES_ele_cld) = traitlist

for (t in 1:length(traitlist)) {
  modeldat <-  comb |> 
    filter(trait == traitlist[t]) |> 
    mutate(elevation = as.factor(elevation), 
           grid = as.factor(grid)) |> 
    drop_na()
  
  
  model<- lme(SES ~ elevation ,
                  random = ~1|grid, 
                  correlation = corSpher(form = ~ x_coord + y_coord|grid, nugget = TRUE), #spherical structure
                  data = modeldat) #only gaussian family possible
  
  em_model <- emmeans(model, specs = "elevation", type = "response")
  comp_letters <-cld(em_model, Letters = letters, adjust = "Tukey", sort = FALSE)

  #Also save results in a list object so we can call it with quarto
  SES_ele_summary[[t]] <- summary(model)
  SES_ele_Rsq[[t]] <- r.squaredGLMM(model)
  SES_ele_anova[[t]] <- anova(model)
  SES_ele_cld[[t]] <- comp_letters
  
}

###Using the logged traits are generally better for model diagnostics than the raw traits
##BUT the leaf area diagnostics are still very bad




###Loop to run SES~microenvironmental variables for all traits####
###=============C5 NULL MODEL, pool = site====================####
###===========================================================###
#import ses
cell_ses_poolsite <- read.csv("All_data/comm_assembly_results/SES_RQ_weighted_cells_C5_poolsite.csv", row.names = 1) |> 
  rename(Cell_ID = cellref) 

#import microenvironmental data
env2 <- read.csv("All_data/clean_data/Environmental data/All_Sites_Environmental_Data.csv") |> 
  #variables we are interested in
  select(Cell_ID:row, rock_cover, northness, soil_moisture_adj_campaign2, 
         soil_depth_CV, mean_soil_depth, slope_height) |> 
  #add elevation variables
  mutate(elevation = case_when(site == "GG" ~ 2000, 
                               site == "WH" ~ 2500, 
                               site == "BK" ~ 3000,
                               .default = NA))

#import remote sensing derived variables
rms <- read.csv("All_data/clean_data/Environmental data/Zonal_stats_all.csv") |> 
  select(CELL_ID, STD) |> 
  rename(Cell_ID = CELL_ID)

#import interpolated microclimate indices
micro_idw <- read.csv("All_data/clean_data/Environmental data/Imke_microclimate_indices_idw_interpolated.csv", row.names = 1)


##Combine SES and environmental data
comb2 <- env2 |> 
  #join to microclimate indices |> 
  full_join(micro_idw, by = "Cell_ID") |> 
  #join to remote sensing data |> 
  full_join(rms, by = "Cell_ID") |> 
  #join, one row in env matches many rows in cell_ses due to it containing ses of different traits
  full_join(cell_ses_poolsite, by = "Cell_ID", relationship = "one-to-many") |>
  mutate(ncolumn = match(column, LETTERS[1:8]), 
         grid = paste0(site, grid)) |> #each grid must have a unique id 
  rename(x_coord = ncolumn, 
         y_coord = row)


###Loop starts here####
#These variables have the best diagnostics after comparing diagnostic plots of log and raw variables
traitlist <- c("Height_cm","Leaf_area_mm2", "log_LDMC","SLA")
sitelist <- c("GG", "WH", "BK")
#lists to store results in
SES_microenv_summary <- vector(mode= "list", length = length(traitlist) * length(sitelist))

SES_microenv_Rsq<- vector(mode= "list", length = length(traitlist) * length(sitelist))

SES_microenv_anova<- vector(mode= "list", length = length(traitlist) * length(sitelist))

l = 1
for (t in 1:length(traitlist)) {
  for(s in 1:length(sitelist)) {
  modeldat <-  comb2 |> 
    filter(trait == traitlist[t], 
           site == sitelist[s]) |> 
    mutate(elevation = as.factor(elevation), 
           grid = as.factor(grid)) |> 
    drop_na()
  
  
  model<- lme(SES ~ mean_T1_growing_season + mean_moist_growing_season + rock_cover+ mean_soil_depth + STD,
              random = ~1|grid, 
              correlation = corSpher(form = ~ x_coord + y_coord|grid, nugget = TRUE), #spherical structure
              data = modeldat) #only gaussian family possible
  
  #Also save results in a list object so we can call it with quarto
  names(SES_microenv_summary)[[l]] <- paste(traitlist[t], sitelist[s], sep = "_")
  SES_microenv_summary[[l]] <- summary(model)
  
  names(SES_microenv_Rsq)[[l]] <- paste(traitlist[t], sitelist[s], sep = "_")
  SES_microenv_Rsq[[l]] <- r.squaredGLMM(model)
  
  names(SES_microenv_anova)[[l]] <- paste(traitlist[t], sitelist[s], sep = "_")
  SES_microenv_anova[[l]] <- anova(model)
  
  l = l+1

}}

```

# Results

`r comb |>  filter(trait == "log_Height") |> drop_na() |> summarise(n = n())` Cells were included in the analysis, with a total species richness of xx. Elevation did not have a consistent effect on the community assembly of different plant traits (Figure 1). Elevation shifted the assembly of plant height and LDMC from environmental filtering at the low elevation, to limiting similarity at the high elevation. Although communities generally became shorter with elevation (Figure 2a), the diversity of plant height was highest at the high elevation (Figure 3), indicating the presence of very short as well as very tall plants in the alpine. There was strong filtering for high LDMC values at the low elevation (indicated by the narrow CWM distribution). This filtering effect was weaker at the mid elevation, and selected for lower LDMC values on average. The assembly of LDMC shifted to trait divergence at the highest elevation, but once again showed on average more conservative strategies than at the mid elevation (Figure 2).

The assembly of leaf area shifted from being no different from random assembly, to environmental filtering at the mid and high elevation (Figure 1). Filtering was in the direction of larger leaves at the mid elevation, and smaller leaves at the high elevation (Figure 2). Environmental filtering was the dominant assembly mechanism at all elevations for SLA, although the filtering effect of SLA was weakest at the mid elevation (Figure 1). SLA was filtered toward conservative leaf-economic strategies at the low and high elevation, but more acquisitive strategies at the mid elevation (Figure 2).

```{r}
#| message: false
#| echo: false
#| label: fig-SES-elevation-ridges
#| fig-cap: "Distributions of standard effect sizes (SES) calculated with plant height, SLA, LDMC and leaf area along an elevational gradient. SES calculated with plant height, LDMC and leaf area used the traits on the log scale."
#| warning: false


l1 <- c("log_Height" = "SES~of~plant~height", "SLA" = "SES~of~SLA", "log_LDMC" = "SES~of~LDMC", "log_LA" = "SES~of~LA")
ridges_letters <- data.frame(trait = c(rep("log_Height", 3), rep("SLA", 3), rep("log_LDMC", 3),rep("log_LA", 3)), 
                             elevation = as.factor(c(rep(c("2000", "2500", "3000"), 4))),
                             letters = c(SES_ele_cld$log_Height$.group, 
                                         SES_ele_cld$SLA$.group, 
                                         SES_ele_cld$log_LDMC$.group, 
                                         SES_ele_cld$log_LA$.group), 
                             x_pos = c(4.5, 4.5, 4.5,
                                       4.5,4.5,4.5, 
                                       9,9,9,
                                       8.1,8.1,8.1), 
                             emmean = c(SES_ele_cld$log_Height$emmean,
                                        SES_ele_cld$SLA$emmean,
                                        SES_ele_cld$log_LDMC$emmean,
                                        SES_ele_cld$log_LA$emmean), 
                             emmean_SE = c(SES_ele_cld$log_Height$SE,
                                           SES_ele_cld$SLA$SE,
                                           SES_ele_cld$log_LDMC$SE,
                                           SES_ele_cld$log_LA$SE))

ses_ridges <- comb |>
  filter(trait %in% c("log_Height", "SLA", "log_LDMC", "log_LA")) |> 
  mutate(elevation = as.factor(elevation)) |> 
  ggplot(aes(x = SES, y = elevation)) +
  geom_density_ridges(alpha = 0.5) +
  facet_wrap(~trait, labeller = as_labeller(l1, default = label_parsed), scale = "free_x", 
             strip.position = "bottom")+
  labs(x = " ", y = "Elevation (m a.s.l.)") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_text(data = ridges_letters, aes(x = x_pos, y = elevation, label = letters, size = 16))+
  geom_point(
    data = ridges_letters,
    aes(x = emmean, y = as.numeric(elevation)),
    size = 2, inherit.aes = FALSE) +
  geom_errorbarh(
    data = ridges_letters,
    aes(xmin = emmean - emmean_SE,
        xmax = emmean + emmean_SE,
        y = as.numeric(elevation)),
    height = 0.08,
    linewidth = 1, inherit.aes = FALSE)+
  theme_bw() +
  theme(legend.position = "none", axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), strip.text = element_text(size = 18), 
        strip.background = element_blank(),
        strip.placement = "outside", panel.grid = element_blank()) 

ses_ridges
```

```{r}
#| label: fig-CWM-elevation-ridges
#| fig-cap: Community-weighted mean traits of cells at the low, mid and high elevation. Height, LDMC and leaf area are on the log scale.
#| warning: false
#| message: false
#| echo: false

#import trait and abundance data
abun_matrix <-read.csv("All_data/comm_assembly_results/abun_matrix.csv", row.names = 1)

mean_traits <- read.csv("All_data/comm_assembly_results/mean_traits.csv", row.names = 1) |> 
  mutate(log_Height = log(Height_cm), 
         log_LA = log(Leaf_area_mm2), 
         log_LDMC = log(LDMC)) |> #use the same log variables as we do in the models
  select(!c(LDMC, Leaf_area_mm2, Height_cm, Thickness_mm))


#compute CWM of each trait for each cell
library(FD)
cwm <- functcomp(x = mean_traits, a = as.matrix(abun_matrix))
cwm2 <- cwm |>
  rownames_to_column(var = "cellref") |> 
  rename(Cell_ID = cellref) |> 
  mutate(elevation = case_when(grepl("BK", Cell_ID) == T ~ "3000", #add elevation variable
                               grepl("WH", Cell_ID) == T ~ "2500",
                               grepl("GG", Cell_ID) == T ~ "2000",.default = NA)) |> 
  pivot_longer(cols = c("log_Height", "log_LDMC", "log_LA", "SLA"), names_to = "trait", values_to = "cwm_value")

#import env data so that we can join x and y coordinates
env <- read.csv("All_data/clean_data/Environmental data/All_Sites_Environmental_Data.csv") |> 
  #variables we are interested in
  select(Cell_ID:row) 

cwm_comb <- env |> 
  #join, one row in env matches many rows in cell_ses due to it containing ses of different traits
  inner_join(cwm2, by = "Cell_ID", relationship = "one-to-many") |>
  mutate(ncolumn = match(column, LETTERS[1:8]), 
         grid = paste0(site, grid)) |> #each grid must have a unique id 
  rename(x_coord = ncolumn, 
         y_coord = row)


####model CWM ~ elevation####
traitlist <- c("log_Height", "log_LDMC", "log_LA", "SLA")
#lists to store results in
CWM_ele_cld<- vector(mode= "list", length = length(traitlist))
names(CWM_ele_cld) = traitlist

for (t in 1:length(traitlist)) {
  modeldat <-  cwm_comb |> 
    filter(trait == traitlist[t]) |> 
    mutate(elevation = as.factor(elevation), 
           grid = as.factor(grid)) |> 
    drop_na()
  
  
  model<- lme(cwm_value ~ elevation ,
                  random = ~1|grid, 
                  correlation = corSpher(form = ~ x_coord + y_coord|grid, nugget = TRUE), #spherical structure
                  data = modeldat) #only gaussian family possible
  
  em_model <- emmeans(model, specs = "elevation", type = "response")
  comp_letters <-cld(em_model, Letters = letters, adjust = "Tukey", sort = FALSE)

  #Also save results in a list object so we can call it with quarto
  CWM_ele_cld[[t]] <- comp_letters
}

##Graph####
l1 <- c("log_Height" = "Plant~height", "SLA" = "SLA", "log_LDMC" = "LDMC", "log_LA" = "LA")
cwm_ridges_letters <- data.frame(trait = c(rep("log_Height", 3), rep("SLA", 3), rep("log_LDMC", 3),rep("log_LA", 3)), 
                             elevation = as.factor(c(rep(c("2000", "2500", "3000"), 4))),
                             letters = c(CWM_ele_cld$log_Height$.group, 
                                         CWM_ele_cld$SLA$.group, 
                                         CWM_ele_cld$log_LDMC$.group, 
                                         CWM_ele_cld$log_LA$.group), 
                             x_pos = c(4,4,4,
                                       62,62,62, 
                                       6.5,6.5,6.5,
                                       8.1,8.1,8.1), 
                             emmean = c(CWM_ele_cld$log_Height$emmean,
                                        CWM_ele_cld$SLA$emmean,
                                        CWM_ele_cld$log_LDMC$emmean,
                                        CWM_ele_cld$log_LA$emmean), 
                             emmean_SE = c(CWM_ele_cld$log_Height$SE,
                                           CWM_ele_cld$SLA$SE,
                                           CWM_ele_cld$log_LDMC$SE,
                                           CWM_ele_cld$log_LA$SE))

cwm_ridges <- cwm2 |>
  filter(trait %in% c("log_Height", "SLA", "log_LDMC", "log_LA")) |> 
  ggplot(aes(x = cwm_value, y = elevation)) +
  geom_density_ridges(alpha = 0.5) +
  facet_wrap(~trait, labeller = as_labeller(l1, default = label_parsed), 
             scale = "free_x", strip.position = "bottom")+
  labs(x = "", y = "Elevation (m a.s.l.)") +
  #scale_x_continuous(breaks = c(0,25,50))+
  theme_bw() +
  geom_text(data = cwm_ridges_letters, aes(x = x_pos, y = elevation, label = letters, size = 16))+
  geom_point(
    data = cwm_ridges_letters,
    aes(x = emmean, y = as.numeric(elevation)),
    size = 2, inherit.aes = FALSE) +
  geom_errorbarh(
    data = cwm_ridges_letters,
    aes(xmin = emmean - emmean_SE,
        xmax = emmean + emmean_SE,
        y = as.numeric(elevation)),
    height = 0.08,
    linewidth = 1, inherit.aes = FALSE)+
  theme(legend.position = "none", axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), strip.text = element_text(size = 18), 
        strip.background = element_blank(),
        strip.placement = "outside", panel.grid = element_blank()) 
cwm_ridges
```

```{r}
#| message: false
#| echo: false
#| label: fig-RaoQ-elevation-ridges
#| fig-cap: "Trait diversity, calculated with RaoQ, of each trait along the elevation gradient"
#| warning: false


####model RaoQ ~ elevation####
traitlist <- c("log_Height", "log_LDMC", "log_LA", "SLA")
#lists to store results in
RaoQ_ele_cld<- vector(mode= "list", length = length(traitlist))
names(RaoQ_ele_cld) = traitlist

for (t in 1:length(traitlist)) {
  modeldat <-  comb |> 
    filter(trait == traitlist[t]) |> 
    mutate(elevation = as.factor(elevation), 
           grid = as.factor(grid)) |> 
    drop_na()
  
  
  model<- lme(RaoQ ~ elevation ,
                  random = ~1|grid, 
                  correlation = corSpher(form = ~ x_coord + y_coord|grid, nugget = TRUE), #spherical structure
                  data = modeldat) #only gaussian family possible
  
  em_model <- emmeans(model, specs = "elevation", type = "response")
  comp_letters <-cld(em_model, Letters = letters, adjust = "Tukey", sort = FALSE)

  #Also save results in a list object so we can call it with quarto
RaoQ_ele_cld[[t]] <- comp_letters
}

###Graph####
l1 <- c("log_Height" = "RaoQ~of~plant~height", "SLA" = "RaoQ~of~SLA", "log_LDMC" = "RaoQ~of~LDMC", "log_LA" = "RaoQ~of~LA")
raoq_ridges_letters <- data.frame(trait = c(rep("log_Height", 3), rep("SLA", 3), rep("log_LDMC", 3),rep("log_LA", 3)), 
                             elevation = as.factor(c(rep(c("2000", "2500", "3000"), 4))),
                             letters = c(RaoQ_ele_cld$log_Height$.group, 
                                         RaoQ_ele_cld$SLA$.group, 
                                         RaoQ_ele_cld$log_LDMC$.group, 
                                         RaoQ_ele_cld$log_LA$.group), 
                             x_pos = c(2.2,2.2,2.2,
                                       2.7,2.7,2.7, 
                                       5,5,5,
                                       3.8,3.8,3.8), 
                             emmean = c(RaoQ_ele_cld$log_Height$emmean,
                                        RaoQ_ele_cld$SLA$emmean,
                                        RaoQ_ele_cld$log_LDMC$emmean,
                                        RaoQ_ele_cld$log_LA$emmean), 
                             emmean_SE = c(RaoQ_ele_cld$log_Height$SE,
                                           RaoQ_ele_cld$SLA$SE,
                                           RaoQ_ele_cld$log_LDMC$SE,
                                           RaoQ_ele_cld$log_LA$SE))

raoq_ridges <- comb |>
  filter(trait %in% c("log_Height", "SLA", "log_LDMC", "log_LA")) |> 
  mutate(elevation = as.factor(elevation)) |> 
  ggplot(aes(x = RaoQ, y = elevation)) +
  geom_density_ridges(alpha = 0.5) +
  facet_wrap(~trait, labeller = as_labeller(l1, default = label_parsed), scale = "free_x", 
             strip.position = "bottom")+
  labs(x = " ", y = "Elevation (m a.s.l.)") +
  theme_bw() +
  geom_text(data = raoq_ridges_letters, aes(x = x_pos, y = elevation, label = letters, size = 16))+
  geom_point(
    data = raoq_ridges_letters,
    aes(x = emmean, y = as.numeric(elevation)),
    size = 2, inherit.aes = FALSE) +
  geom_errorbarh(
    data = raoq_ridges_letters,
    aes(xmin = emmean - emmean_SE,
        xmax = emmean + emmean_SE,
        y = as.numeric(elevation)),
    height = 0.08,
    linewidth = 1, inherit.aes = FALSE)+
  theme(legend.position = "none", axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), strip.text = element_text(size = 18), 
        strip.background = element_blank(),
        strip.placement = "outside", panel.grid = element_blank()) 

raoq_ridges
```

Micro-environmental factors were related to fine-scale variation in community assembly, although the relationships varied greatly between traits and sites. The models had low explanatory power, with R squared values generally below 0.1 (Table 1).

At the low site, soil temperature filtered size traits (height and leaf area) to shorter plants and smaller leaves (Table 2, Figure 4a,b). Surface heterogeneity was related to divergence of height, while soil moisture and soil depth were related to divergence in leaf area (Table 2). At the mid site, rock cover was related to divergence of height and LDMC, while soil depth and surface heterogeneity was related to divergence in leaf area (Table 3). Soil moisture was related to filtering of LDMC (Table 3), although this filtering effect was not strongly directional (LDMC did not change with soil moisture, Figure 4d). Soil depth was related to the filtering of SLA toward more acquisitive leaf-economic strategies (Table 3, Figure 4c). At the high site, soil moisture and surface heterogeneity was related to divergence in plant height, and rock cover was related to divergence of leaf area (Table 4). Soil depth had a filtering effect on leaf area, but was not strongly directional (Figure 4e). Surface heterogeneity was related to divergence of LDMC and SLA (Table 4). Divergence of SLA at the high site was also related to rock cover and soil temperature (Table 4).

Surface heterogeneity and rock cover were always related to divergence in size and leaf-economic traits. The directions of the effects of other micro-environmental variables on assembly were not consistent across sites.

```{r}
#| label: tbl-R2-models
#| tbl-cap: Marginal and conditional R2 values of micro-environmental models in predicting SES of each trait at the low, mid and high site.
#| echo: false
#| message: false
#| warning: false

library(flextable)

R2_table <- tibble(trait = c("Height", "Leaf area", "LDMC", "SLA"), 
                   GG_R2M = c(SES_microenv_Rsq$Height_cm_GG[1,][1], SES_microenv_Rsq$Leaf_area_mm2_GG[1,][1], SES_microenv_Rsq$log_LDMC_GG[1,][1], SES_microenv_Rsq$SLA_GG[1,][1]), 
                   GG_R2C = c(SES_microenv_Rsq$Height_cm_GG[1,][2], SES_microenv_Rsq$Leaf_area_mm2_GG[1,][2], SES_microenv_Rsq$log_LDMC_GG[1,][2], SES_microenv_Rsq$SLA_GG[1,][2]), 
                   WH_R2M = c(SES_microenv_Rsq$Height_cm_WH[1,][1], SES_microenv_Rsq$Leaf_area_mm2_WH[1,][1], SES_microenv_Rsq$log_LDMC_WH[1,][1], SES_microenv_Rsq$SLA_WH[1,][1]), 
                   WH_R2C = c(SES_microenv_Rsq$Height_cm_WH[1,][2], SES_microenv_Rsq$Leaf_area_mm2_WH[1,][2], SES_microenv_Rsq$log_LDMC_WH[1,][2], SES_microenv_Rsq$SLA_WH[1,][2]),
                   BK_R2M = c(SES_microenv_Rsq$Height_cm_BK[1,][1], SES_microenv_Rsq$Leaf_area_mm2_BK[1,][1], SES_microenv_Rsq$log_LDMC_BK[1,][1], SES_microenv_Rsq$SLA_BK[1,][1]), 
                   BK_R2C = c(SES_microenv_Rsq$Height_cm_BK[1,][2], SES_microenv_Rsq$Leaf_area_mm2_BK[1,][2], SES_microenv_Rsq$log_LDMC_BK[1,][2], SES_microenv_Rsq$SLA_BK[1,][2]))

R2_table |>
  flextable() |>
  colformat_double(j = 2:7, digits = 3) |>
  set_header_labels(
    trait  = "Trait",
    GG_R2M = "Marginal R\u00B2",
    GG_R2C = "Conditional R\u00B2",
    WH_R2M = "Marginal R\u00B2",
    WH_R2C = "Conditional R\u00B2",
    BK_R2M = "Marginal R\u00B2",
    BK_R2C = "Conditional R\u00B2"
  ) |>
  add_header_row(
    values    = c("Trait", "Low", "Mid", "High"),
    colwidths = c(1, 2, 2, 2),
    top       = TRUE
  ) |>
  merge_v(j = 1, part = "header") |>
  align(align = "center", part = "header") |>
  align(align = "center", part = "body", j = 2:7) |>
  valign(valign = "bottom", part = "header") |>
  theme_vanilla() |>
  autofit() |>
  fit_to_width(max_width = 6.5)

```

```{r}
#| label: tbl-GG-microenv
#| tbl-cap: Effects of microenvironmental variables on assembly of traits at the low site. Coefficients show the effect of microenvironmental variables on the SES calculated with each trait.
#| echo: false
#| warning: false


GG_microenv_reults <- tibble(variable = c("mean soil temperature", "mean soil moisture", "rock cover", "mean soil     depth", "surface heterogeneity"), 
                height_coef = SES_microenv_summary$Height_cm_GG$coefficients$fixed[2:6], 
                height_pval = SES_microenv_anova$Height_cm_GG$`p-value`[2:6], 
                la_coef = SES_microenv_summary$Leaf_area_mm2_GG$coefficients$fixed[2:6], 
                la_pval = SES_microenv_anova$Leaf_area_mm2_GG$`p-value`[2:6],
                ldmc_coef = SES_microenv_summary$log_LDMC_GG$coefficients$fixed[2:6], 
                ldmc_pval = SES_microenv_anova$log_LDMC_GG$`p-value`[2:6],
                sla_coef = SES_microenv_summary$SLA_GG$coefficients$fixed[2:6], 
                sla_pval = SES_microenv_anova$SLA_GG$`p-value`[2:6])

Rsq <- c(SES_microenv_Rsq$Height_cm_GG[1,][1], SES_microenv_Rsq$Leaf_area_mm2_GG[1,][1], 
                 SES_microenv_Rsq$log_LDMC_GG[1,][1], SES_microenv_Rsq$SLA_GG[1,][1])

GG_microenv_reults |> 
  gt() |> 
  cols_label(height_coef = "coefficient", 
             height_pval = "p-value", 
             la_coef = "coefficient", 
             la_pval = "p-value", 
             ldmc_coef = "coefficient", 
             ldmc_pval = "p-value", 
             sla_coef = "coefficient", 
             sla_pval = "p-value") |> 
  fmt_number(columns = everything(), decimals = 3) |> 
  tab_spanner(label = "Height", columns = c(height_coef, height_pval)) |>  
  tab_spanner(label = "Leaf area", columns = c(la_coef, la_pval)) |>
  tab_spanner(label = "LDMC", columns = c(ldmc_coef, ldmc_pval)) |>
  tab_spanner(label = "SLA", columns = c(sla_coef, sla_pval)) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = height_pval, rows = height_pval < 0.05)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = la_pval, rows = la_pval < 0.05)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = ldmc_pval, rows = ldmc_pval < 0.05)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = sla_pval, rows = sla_pval < 0.05)
  ) |> 
  tab_options(table.width = pct(100))

```

```{r}
#| label: tbl-WH-microenv
#| tbl-cap: Effects of microenvironmental variables on assembly of traits at the mid site. Coefficients show the effect of microenvironmental variables on the SES calculated with each trait.
#| echo: false
#| warning: false

WH_microenv_reults <- tibble(variable = c("mean soil temperature", "mean soil moisture", "rock cover", "mean soil     depth", "surface heterogeneity"), 
                height_coef = SES_microenv_summary$Height_cm_WH$coefficients$fixed[2:6], 
                height_pval = SES_microenv_anova$Height_cm_WH$`p-value`[2:6], 
                la_coef = SES_microenv_summary$Leaf_area_mm2_WH$coefficients$fixed[2:6], 
                la_pval = SES_microenv_anova$Leaf_area_mm2_WH$`p-value`[2:6],
                ldmc_coef = SES_microenv_summary$log_LDMC_WH$coefficients$fixed[2:6], 
                ldmc_pval = SES_microenv_anova$log_LDMC_WH$`p-value`[2:6],
                sla_coef = SES_microenv_summary$SLA_WH$coefficients$fixed[2:6], 
                sla_pval = SES_microenv_anova$SLA_WH$`p-value`[2:6])

Rsq <- c(SES_microenv_Rsq$Height_cm_WH[1,][1], SES_microenv_Rsq$Leaf_area_mm2_WH[1,][1], 
                 SES_microenv_Rsq$log_LDMC_WH[1,][1], SES_microenv_Rsq$SLA_WH[1,][1])

WH_microenv_reults |> 
  gt() |> 
  cols_label(height_coef = "coefficient", 
             height_pval = "p-value", 
             la_coef = "coefficient", 
             la_pval = "p-value", 
             ldmc_coef = "coefficient", 
             ldmc_pval = "p-value", 
             sla_coef = "coefficient", 
             sla_pval = "p-value") |> 
  fmt_number(columns = everything(), decimals = 3) |> 
  tab_spanner(label = "Height", columns = c(height_coef, height_pval)) |>  
  tab_spanner(label = "Leaf area", columns = c(la_coef, la_pval)) |>
  tab_spanner(label = "LDMC", columns = c(ldmc_coef, ldmc_pval)) |>
  tab_spanner(label = "SLA", columns = c(sla_coef, sla_pval)) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = height_pval, rows = height_pval < 0.05)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = la_pval, rows = la_pval < 0.05)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = ldmc_pval, rows = ldmc_pval < 0.05)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = sla_pval, rows = sla_pval < 0.05)
  ) |> 
  tab_options(table.width = pct(100))

```

```{r}
#| label: tbl-BK-microenv
#| tbl-cap: Effects of microenvironmental variables on assembly of traits at the high site. Coefficients show the effect of microenvironmental variables on the SES calculated with each trait.
#| echo: false
#| warning: false

BK_microenv_reults <- tibble(variable = c("mean soil temperature", "mean soil moisture", "rock cover", "mean soil     depth", "surface heterogeneity"), 
                height_coef = SES_microenv_summary$Height_cm_BK$coefficients$fixed[2:6], 
                height_pval = SES_microenv_anova$Height_cm_BK$`p-value`[2:6], 
                la_coef = SES_microenv_summary$Leaf_area_mm2_BK$coefficients$fixed[2:6], 
                la_pval = SES_microenv_anova$Leaf_area_mm2_BK$`p-value`[2:6],
                ldmc_coef = SES_microenv_summary$log_LDMC_BK$coefficients$fixed[2:6], 
                ldmc_pval = SES_microenv_anova$log_LDMC_BK$`p-value`[2:6],
                sla_coef = SES_microenv_summary$SLA_BK$coefficients$fixed[2:6], 
                sla_pval = SES_microenv_anova$SLA_BK$`p-value`[2:6])

Rsq <- c(SES_microenv_Rsq$Height_cm_BK[1,][1], SES_microenv_Rsq$Leaf_area_mm2_BK[1,][1], 
                 SES_microenv_Rsq$log_LDMC_BK[1,][1], SES_microenv_Rsq$SLA_BK[1,][1])

BK_microenv_reults |> 
  gt() |> 
  cols_label(height_coef = "coefficient", 
             height_pval = "p-value", 
             la_coef = "coefficient", 
             la_pval = "p-value", 
             ldmc_coef = "coefficient", 
             ldmc_pval = "p-value", 
             sla_coef = "coefficient", 
             sla_pval = "p-value") |> 
  fmt_number(columns = everything(), decimals = 3) |> 
  tab_spanner(label = "Height", columns = c(height_coef, height_pval)) |>  
  tab_spanner(label = "Leaf area", columns = c(la_coef, la_pval)) |>
  tab_spanner(label = "LDMC", columns = c(ldmc_coef, ldmc_pval)) |>
  tab_spanner(label = "SLA", columns = c(sla_coef, sla_pval)) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = height_pval, rows = height_pval < 0.05)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = la_pval, rows = la_pval < 0.05)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = ldmc_pval, rows = ldmc_pval < 0.05)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = sla_pval, rows = sla_pval < 0.05)
  ) |> 
  tab_options(table.width = pct(100))

```

```{r cwm microenvironment models and scatterplots}
#| label: fig-CWM-microenvironmental-scatterplots
#| fig-cap: Direction of environmental filtering by micro-environmental variables on community-weighted mean traits at the low (a,b), mid site (c) and high elevation (d). LDMC is shown on the log scale.
#| echo: false
#| message: false
#| warning: false


#import trait and abundance data
abun_matrix <-read.csv("All_data/comm_assembly_results/abun_matrix.csv", row.names = 1)

mean_traits <- read.csv("All_data/comm_assembly_results/mean_traits.csv", row.names = 1) |> 
  mutate(log_Height = log(Height_cm), 
         log_LA = log(Leaf_area_mm2), 
         log_LDMC = log(LDMC))


#compute CWM of each trait for each cell
library(FD)
cwm <- functcomp(x = mean_traits, a = as.matrix(abun_matrix))
cwm2 <- cwm |>
  rownames_to_column(var = "cellref") |> 
  rename(Cell_ID = cellref) |> 
  pivot_longer(cols = !Cell_ID, names_to = "trait", values_to = "cwm_value")

#import microenvironmental data
env2 <- read.csv("All_data/clean_data/Environmental data/All_Sites_Environmental_Data.csv") |> 
  #variables we are interested in
  select(Cell_ID:row, rock_cover, northness, soil_moisture_adj_campaign2, 
         soil_depth_CV, mean_soil_depth, slope_height) |> 
  mutate(mean_soil_depth = case_when(mean_soil_depth > 61 ~ 61, 
                                     .default = mean_soil_depth)) |> 
  #add elevation variables
  mutate(elevation = case_when(site == "GG" ~ 2000, 
                               site == "WH" ~ 2500, 
                               site == "BK" ~ 3000,
                               .default = NA))

#import remote sensing derived variables
rms <- read.csv("All_data/clean_data/Environmental data/Zonal_stats_all.csv") |> 
  select(CELL_ID, STD) |> 
  rename(Cell_ID = CELL_ID)

#import interpolated microclimate indices
micro_idw <- read.csv("All_data/clean_data/Environmental data/Imke_microclimate_indices_idw_interpolated.csv", row.names = 1)


##Combine CWM and environmental data
cwm_comb <- env2 |> 
  #join to microclimate indices |> 
  full_join(micro_idw, by = "Cell_ID") |> 
  #join to remote sensing data |> 
  full_join(rms, by = "Cell_ID") |> 
  #join, one row in env matches many rows in cell_ses due to it containing ses of different traits
  full_join(cwm2, by = "Cell_ID", relationship = "one-to-many") |>
  mutate(ncolumn = match(column, LETTERS[1:8]), 
         grid = paste0(site, grid)) |> #each grid must have a unique id 
  rename(x_coord = ncolumn, 
         y_coord = row)


####model CWM ~ microenvironmental variables####
traitlist <- c("Height_cm","Leaf_area_mm2", "log_LDMC","SLA") #the traits we chose to log here are different to the traits logged in the elevation model.
sitelist <- c("GG", "WH", "BK")
variables <- c("mean_T1_growing_season" , "mean_moist_growing_season" , "rock_cover", "mean_soil_depth" , "STD")
#list to save predictions in 
predictions_list <- vector(mode= "list", length = length(traitlist)*length(sitelist)*length(variables))

l = 1
for (t in 1:length(traitlist)) {
  for(s in 1:length(sitelist)) {
  modeldat <-  cwm_comb |> 
    filter(trait == traitlist[t], 
           site == sitelist[s]) |> 
    mutate(elevation = as.factor(elevation), 
           grid = as.factor(grid)) |> 
    drop_na()
  
  
  model<- lm(cwm_value ~ mean_T1_growing_season + mean_moist_growing_season + rock_cover+ mean_soil_depth + STD,
              data = modeldat) 
  
  #Generate model predictions for each variable setting every other variable to its mean
  for(v in 1:length(variables)) {
    pred_var <- variables[v]
    mean_var <- setdiff(variables, variables[v])
    
    pred_dat <- modeldat |> 
      mutate(across(all_of(mean_var), ~ mean(.x, na.rm = TRUE))) 
    
    predicted_cwm <- predict(model, newdata = pred_dat, type = "response", se.fit = TRUE)
    fitted <- predicted_cwm$fit
    se <- predicted_cwm$se.fit
    
    plot_data <- pred_dat |> 
      select(all_of(variables)) |> 
      mutate(predicted_cwm = fitted, 
             standard_error = se) 
    
    predictions_list[[l]] <- plot_data
    names(predictions_list)[[l]] <- paste("CWM", traitlist[t], sitelist[s], "model", variables[v], "predictions", sep = "_")
  
    l = l+1
    }
  }}


####Plotting predictions
df1 <- predictions_list$CWM_Height_cm_GG_model_mean_T1_growing_season_predictions
H_temp_GG <- cwm_comb |> 
  filter(trait == "Height_cm", site == "GG") |> 
  ggplot(aes(x = mean_T1_growing_season, y = cwm_value)) +
  geom_point() +
  geom_line(data = df1, aes(x = mean_T1_growing_season, y = predicted_cwm), color= "red", size = 1) +
  geom_ribbon(data = df1, aes(x = mean_T1_growing_season, y = predicted_cwm, 
                              ymin = predicted_cwm - standard_error, ymax = predicted_cwm + standard_error), fill = "red", alpha = 0.2)+
  labs(x = "Mean soil temperature", y = "CWM Height") +
  theme_classic()

df2 <- predictions_list$CWM_Leaf_area_mm2_GG_model_mean_T1_growing_season_predictions
LA_temp_GG <-  cwm_comb |> 
  filter(trait == "Leaf_area_mm2", site == "GG") |> 
  ggplot(aes(x = mean_T1_growing_season, y = cwm_value)) +
  geom_point() +
  geom_line(data = df2, aes(x = mean_T1_growing_season, y = predicted_cwm), color= "red", size = 1) +
  geom_ribbon(data = df2, aes(x = mean_T1_growing_season, y = predicted_cwm, 
                              ymin = predicted_cwm - standard_error, ymax = predicted_cwm + standard_error), fill = "red", alpha = 0.2)+
  labs(x = "Mean soil temperature", y = "CWM Leaf area") +
  theme_classic()

df3 <- predictions_list$CWM_SLA_WH_model_mean_soil_depth_predictions
SLA_sdepth_WH <-  cwm_comb |> 
  filter(trait == "SLA", site == "WH") |> 
  ggplot(aes(x = mean_soil_depth, y = cwm_value)) +
  geom_point() +
  geom_line(data = df3, aes(x = mean_soil_depth, y = predicted_cwm), color= "red", size = 1) +
  geom_ribbon(data = df3, aes(x = mean_soil_depth, y = predicted_cwm, 
                              ymin = predicted_cwm - standard_error, ymax = predicted_cwm + standard_error), fill = "red", alpha = 0.2)+
  labs(x = "Mean soil depth", y = "CWM SLA") +
  theme_classic()

df5 <- predictions_list$CWM_log_LDMC_WH_model_mean_moist_growing_season_predictions
LDMC_smoist_WH <-  cwm_comb |> 
  filter(trait == "log_LDMC", site == "WH") |> 
  ggplot(aes(x = mean_moist_growing_season, y = cwm_value)) +
  geom_point() +
  geom_line(data = df5, aes(x = mean_moist_growing_season, y = predicted_cwm), color= "red", size = 1) +
  geom_ribbon(data = df5, aes(x = mean_moist_growing_season, y = predicted_cwm, 
                              ymin = predicted_cwm - standard_error, ymax = predicted_cwm + standard_error), fill = "red", alpha = 0.2) +
  labs(x = "Mean soil moisture", y = "CWM LDMC") +
  theme_classic()

df4 <- predictions_list$CWM_Leaf_area_mm2_BK_model_mean_soil_depth_predictions
LA_sdepth_BK <-  cwm_comb |> 
  filter(trait == "Leaf_area_mm2", site == "BK") |> 
  ggplot(aes(x = mean_soil_depth, y = cwm_value)) +
  geom_point() +
  geom_line(data = df4, aes(x = mean_soil_depth, y = predicted_cwm), color= "red", size = 1) +
  geom_ribbon(data = df4, aes(x = mean_soil_depth, y = predicted_cwm, 
                              ymin = predicted_cwm - standard_error, ymax = predicted_cwm + standard_error), fill = "red", alpha = 0.2)+
  labs(x = "Mean soil depth", y = "CWM Leaf area") +
  theme_classic()


combined_fig <- ggarrange(H_temp_GG, LA_temp_GG, SLA_sdepth_WH,LDMC_smoist_WH, LA_sdepth_BK, labels = "auto")
combined_fig

```
