---
title: "Ch2_results"
format: html
editor: visual
---

```{r setup chunk}
#| message: false
#| echo: false
#| warning: false
library(tidyverse)
library(ggplot2)
library(ggridges)
library(nlme)
library(emmeans)
library(multcomp)
library(MuMIn)
library(conflicted)
conflict_prefer_all("dplyr", quiet = TRUE)

#import SES data
cell_ses <- read.csv("All_data/comm_assembly_results/SES_RQ_weighted_cells_C5_entire.csv", row.names = 1) |> 
  rename(Cell_ID = cellref)

#import microenvironmental data
env <- read.csv("All_data/clean_data/Environmental data/All_Sites_Environmental_Data.csv") |> 
  #variables we are interested in
  select(Cell_ID:row, rock_cover, northness, soil_moisture_adj_campaign2, 
           soil_depth_CV, mean_soil_depth, slope_height) |> 
  #add elevation variables
  mutate(elevation = case_when(site == "GG" ~ 2000, 
                               site == "WH" ~ 2500, 
                               site == "BK" ~ 3000,
                               .default = NA))

#import remote sensing derived variables
rms <- read.csv("All_data/clean_data/Environmental data/Zonal_stats_all.csv") |> 
  select(CELL_ID, STD) |> 
  rename(Cell_ID = CELL_ID)

#import interpolated microclimate indices
micro_idw <- read.csv("All_data/clean_data/Environmental data/Imke_microclimate_indices_idw_interpolated.csv", row.names = 1)


##Combine SES and environmental data
comb <- env |> 
  #join to microclimate indices |> 
  full_join(micro_idw, by = "Cell_ID") |> 
  #join to remote sensing data |> 
  full_join(rms, by = "Cell_ID") |> 
  #join, one row in env matches many rows in cell_ses due to it containing ses of different traits
  full_join(cell_ses, by = "Cell_ID", relationship = "one-to-many") |>
  mutate(ncolumn = match(column, LETTERS[1:8]), 
         grid = paste0(site, grid)) |> #each grid must have a unique id 
  rename(x_coord = ncolumn, 
         y_coord = row)


###=========================================###
###Loop to run SES~elevation for all traits####
###============C5 NULL MODEL===============####
###=========================================###
traitlist <- c("log_Height", "log_SLA", "log_LDMC", "log_LA", "Height_cm", "SLA", "LDMC", "Leaf_area_mm2")
#lists to store results in
SES_ele_summary <- vector(mode= "list", length = length(traitlist))
names(SES_ele_summary) = traitlist

SES_ele_Rsq<- vector(mode= "list", length = length(traitlist))
names(SES_ele_Rsq) = traitlist

SES_ele_anova<- vector(mode= "list", length = length(traitlist))
names(SES_ele_anova) = traitlist

SES_ele_cld<- vector(mode= "list", length = length(traitlist))
names(SES_ele_cld) = traitlist

for (t in 1:length(traitlist)) {
  modeldat <-  comb |> 
    filter(trait == traitlist[t]) |> 
    mutate(elevation = as.factor(elevation), 
           grid = as.factor(grid)) |> 
    drop_na()
  
  
  model<- lme(SES ~ elevation ,
                  random = ~1|grid, 
                  correlation = corSpher(form = ~ x_coord + y_coord|grid, nugget = TRUE), #spherical structure
                  data = modeldat) #only gaussian family possible
  
  em_model <- emmeans(model, specs = "elevation", type = "response")
  comp_letters <-cld(em_model, Letters = letters, adjust = "Tukey", sort = FALSE)
  
  #Also save results in a list object so we can call it with quarto
  SES_ele_summary[[t]] <- summary(model)
  SES_ele_Rsq[[t]] <- r.squaredGLMM(model)
  SES_ele_anova[[t]] <- anova(model)
  SES_ele_cld[[t]] <- comp_letters
  
}

```

# Results

`r comb |>  filter(trait == "log_Height") |> drop_na() |> summarise(n = n())` Cells were included in the analysis, with a total species richness of xx. Elevation did not have a consistent effect on the community assembly of different plant traits (Figure 1). Elevation shifted the assembly of plant height and LDMC from environmental filtering at the low elevation, to limiting similarity at the high elevation. Environmental filtering was the dominant assembly mechanism at all elevations for leaf area and SLA, although the filtering effect of LSA was strongest at the mid elevation.

```{r}
#| message: false
#| echo: false
#| label: fig-SES-elevation-ridges
#| fig-cap: "The deviation of trait diversity calculated with Height, SLA, LDMC and leaf area from random assembly (SES)"
#| warning: false


l1 <- c("log_Height" = "SES~of~plant~height", "log_SLA" = "SES~of~SLA", "log_LDMC" = "SES~of~LDMC", "log_LA" = "SES~of~LA")
ridges_letters <- data.frame(trait = c(rep("log_Height", 3), rep("log_SLA", 3), rep("log_LDMC", 3),rep("log_LA", 3)), 
                             elevation = as.factor(c(rep(c("2000", "2500", "3000"), 4))),
                             letters = c(SES_ele_cld$log_Height$.group, 
                                         SES_ele_cld$log_SLA$.group, 
                                         SES_ele_cld$log_LDMC$.group, 
                                         SES_ele_cld$log_LA$.group), 
                             x_pos = c(4.5, 4.5, 4.5,
                                       6,6,6, 
                                       9,9,9,
                                       8.1,8.1,8.1), 
                             emmean = c(SES_ele_cld$log_Height$emmean,
                                        SES_ele_cld$log_SLA$emmean,
                                        SES_ele_cld$log_LDMC$emmean,
                                        SES_ele_cld$log_LA$emmean), 
                             emmean_SE = c(SES_ele_cld$log_Height$SE,
                                           SES_ele_cld$log_SLA$SE,
                                           SES_ele_cld$log_LDMC$SE,
                                           SES_ele_cld$log_LA$SE))

ses_ridges <- comb |>
  filter(trait %in% c("log_Height", "log_SLA", "log_LDMC", "log_LA")) |> 
  mutate(elevation = as.factor(elevation)) |> 
  ggplot(aes(x = SES, y = elevation)) +
  geom_density_ridges(alpha = 0.5) +
  facet_wrap(~trait, labeller = as_labeller(l1, default = label_parsed), scale = "free_x", 
             strip.position = "bottom")+
  labs(x = " ", y = "Elevation (m a.s.l.)") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_text(data = ridges_letters, aes(x = x_pos, y = elevation, label = letters, size = 16))+
  geom_point(
    data = ridges_letters,
    aes(x = emmean, y = as.numeric(elevation)),
    size = 2, inherit.aes = FALSE) +
  geom_errorbarh(
    data = ridges_letters,
    aes(xmin = emmean - emmean_SE,
        xmax = emmean + emmean_SE,
        y = as.numeric(elevation)),
    height = 0.08,
    linewidth = 1, inherit.aes = FALSE)+
  theme_bw() +
  theme(legend.position = "none", axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), strip.text = element_text(size = 18), 
        strip.background = element_blank(),
        strip.placement = "outside", panel.grid = element_blank()) 

ses_ridges
```
